Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2015-11-08 draft.tex
replace 'condition' with 'restriction'

The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors. That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$. Strategies for the patron and the international community are how much to invest in each of the six state variables. However, positive investments in some of the state variables can be ruled out by preference assumptions.

@@

The restrictions on the class of games in $\mathcal{G}$ are sufficient but not necessary. For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Condition (6) need not be met to maintain the status quo in the short run. The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time. In cases where condition (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.

@@

To begin, note that either outside actor could invest toward increasing any of the six state variables in a period: $q_{st},q_{gt},w_{st},w_{gt},l_{st},$ and $l_{gt}$.\footnote{In the case of $w_{st}$ and $w_{gt}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} It is not necessary to focus on which investment vehicle is optimal for a government to choose; what is essential is to determine which outcomes---i.e. \emph{Recognition, Status Quo}, or \emph{Reunification}---they will target given that they will choose the most efficient way to alter the payoffs of players $s$ and $g$.

%Given their preference orderings, some of these can be ruled out. Player $p$ will not invest to increase $s$'s payoffs from ceding ($l_{sn}$) or $g$'s payoffs from war when it could do so for $s$. It will also not invest in $q_{gn}$ when it can instead increase $l_{gn}$, leaving the list of rational investment choices for $p$ as $q_{sn},l_{gn}$, and $w_{sn}$. Symmetrically, $c$ will not invest in $q_{sn},L_{gn}$, or $w_{sn}$. Given the assumption that the international community follows a norm of encouraging peaceful resolution, it also will not invest in $w_{gn}$, leaving only $l_{sn}$ and $q_{gn}$ as rational investment options for player $c$.

@@

\begin{lemma}
When Condition 3 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
\end{lemma}

\noindent Condition 3 provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome. To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}

Consider period 1. When Condition (4) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest. That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment. Thus we need Condition (5): $B_{p1} \geq\frac{\beta}{1-\de} - \left(q_{s1} - l_{s1} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $t > 1$, there are two cases to consider. Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$. Conditions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$. However, because in each period $q_{st}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Condition (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Condition 6).

@@

\begin{lemma}
When Condition 4 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
\end{lemma}

In the equilibrium under consideration here, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome. This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved. The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Conditions (5) and (6): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}

@@

\begin{lemma}
Conditions 3 and 4 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
\end{lemma}

@@

The conditions of Proposition 1 \emph{do not} provide for a unique equilibrium, or even a unique equilibrium outcome. In fact, at least one additional equilibrium outcome always coexists with the status quo outcome.

@@

\begin{proposition}
Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo. In order for sanctions to lead to ceding by the secessionists, the following are required:

@@

If condition (1) fails, player $p$ will continue to invest to prevent reunification as in Proposition 1. If conditions (2) or (3) fail, sanctions will lead to fighting initiated by the secessionists---either supported by the patron, or without its support in the case of condition 3. Note here from condition (2) that sanctions can induce investment behavior by the patron that was ruled out under the conditions of Proposition 1: the goal of sanctions is to destabilize the Status Quo equilibrium and they certainly can achieve that goal but there may be unintended consequences, most notably the initiation of war by the secessionists.

>>>>

The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors. That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.\footnote{In the case of $w_{st}$ and $w_{gt}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables. However, positive investments in some of the state variables can be ruled out by preference assumptions.

@@

The restrictions on the class of games in $\mathcal{G}$ are sufficient but not necessary. For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (6) need not be met to maintain the status quo in the short run. The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time. In cases where Restriction (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.

@@

Recall that either outside actor could invest toward increasing any of the six state variables in a period: $q_{st},q_{gt},w_{st},w_{gt},l_{st},$ and $l_{gt}$. However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2. The remainder of the proof treats each set of matched investment decisions in turn.

@@

\begin{lemma}
When Restriction (3) of Definition 1 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
\end{lemma}

\noindent Restriction (3) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome. To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}

Consider period 1. When Restriction (4) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest. That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment. Thus we need Restriction (5): $B_{p1} \geq\frac{\beta}{1-\de} - \left(q_{s1} - l_{s1} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $t > 1$, there are two cases to consider. Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$. Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$. However, because in each period $q_{st}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).

@@

\begin{lemma}
When Restriction 4 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
\end{lemma}

In the equilibrium under consideration here, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome. This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved. The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Restrictions (5) and (6): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}

@@

\begin{lemma}
Restrictions 3 and 4 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
\end{lemma}

@@

The restrictions in Definition 1 \emph{do not} provide for a unique equilibrium, or even a unique equilibrium outcome. In fact, at least one additional equilibrium outcome always coexists with the status quo outcome.

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo. In order for sanctions to lead to ceding by the secessionists, the following are required:

@@

If Restriction (1) fails, player $p$ will continue to invest to prevent reunification as in Proposition 1. If Restrictions (2) or (3) fail, sanctions will lead to fighting initiated by the secessionists---either supported by the patron, or without its support in the case of Restriction 3. Note here from Restriction (2) that sanctions can induce investment behavior by the patron that was ruled out under the Restrictions of Definition 1: the goal of sanctions is to destabilize the Status Quo equilibrium and they certainly can achieve that goal but there may be unintended consequences, most notably the initiation of war by the secessionists.
